Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-05-28 writeup/gz_hubble_data.tex
Science motivation for studying bars. Closes #128.

In addition to the physical parameters of the input images, the \ferengi{}
output depends on assumptions of the global galaxy evolution model. This
evolution is a crude mechanism that mimics the brightness increase of galaxies
with increasing redshift \citep[\eg,][]{lil98,lov12}. The effect on the
redshifted images is simply an empirical addition to the magnitude of a galaxy
of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and
$e$ is the evolutionary correction in magnitudes ($e=-1$ essentially
brightens the entire galaxy by 1~magnitude by $z=1$). \ferengi{} was run on the images
for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments
of $\Delta e = 0.5$. Figure~\ref{fig:exampleFERENGI} shows several examples
of the effects of ``losing'' spiral/disc features with increasing redshift
for two galaxies with no evolution corrections ($e=0$); as the signal-to-noise ratio in the images
decreases and the galaxies become fainter, the contrast between features
like spiral arms goes down and \ffeatures{} drops to the point where both
galaxies would have been classified as likely ellipticals.

@@

A sample of barred disk galaxies is created by applying cuts on the previous
tasks as listed in Table~\ref{tbl:thresholds}. We identify 11,049 ``bar
candidates'' as disk galaxies for which bar identification can be attempted in
GZH. These galaxies were selected by applying the cuts $N_{bar} \ge 20$,
$f_{\rm features}>0.23, f_{\rm clumpy,no} > 0.30,$ and $f_{edgeon,no}>0.25$.
These galaxies are featured, non-clumpy, non-edge~on galaxies. Of these, a pure
sample of 730~barred disks was identified by applying a cut of \fbar$>0.7$. A
subsample of 50~galaxies were visually inspected and 94\% were found to contain
strong bars. A complete sample of strong and weak bars was created by applying
a cut of \fbar$>0.3$.  This sample contained 3,218~galaxies, 86\% of which were
found to contain weak or strong bars through visual inspection.

The resulting bar sample can be used to estimate the redshift evolution of bar
fraction; we find a steady decrease of \fbar$\sim 0.32$ at $z=0.4$ to
\fbar$\sim 0.24$ at $z=1.0$. This decrease agrees with
\citet{mel14}, although they report a lower overall bar fraction going from
\fbar$=0.22$ at $z=0.4$ to \fbar$=0.11$ at $z=1.0$. The difference in total bar
fraction is expected, as this analysis used a looser cut on \fbar, there is no
luminosity cut, and the use of debiased values for \ffeatures{} increases the
total amount of disks in the sample compared to \citet{mel14}. However, the
results from GZH indicate that if features are reliably identified, the bar is
always visible; for galaxies at increasing redshifts, the discs fade first but
the galactic bar remains visible \citep{mel16}. As a result, we do \emph{not}
recommend using an evolving cut on \fbar{} as a function of redshift for the
selection of barred discs.

>>>>

In addition to the physical parameters of the input images, the \ferengi{}
output depends on assumptions of the global galaxy evolution model. This
evolution is a crude mechanism that mimics the brightness increase of galaxies
with increasing redshift \citep[\eg,][]{lil98,lov12}. The effect on the
redshifted images is simply an empirical addition to the magnitude of a galaxy
of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and
$e$ is the evolutionary correction in magnitudes ($e=-1$ essentially
brightens the entire galaxy by 1~magnitude by $z=1$). \ferengi{} was run on the images
for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments
of $\Delta e = 0.5$. Figure~\ref{fig:exampleFERENGI} shows several examples
of the effects of ``losing'' spiral/disk features with increasing redshift
for two galaxies with no evolution corrections ($e=0$); as the signal-to-noise ratio in the images
decreases and the galaxies become fainter, the contrast between features
like spiral arms goes down and \ffeatures{} drops to the point where both
galaxies would have been classified as likely ellipticals.

@@

Bars in galaxies are a trace of the dynamical state of the disk
\citep[\eg,][]{com09a,ath12}. Disks which have significant vertical motions
typically do not form bars (with the exception of tidally triggered bars;
\citealt{bar91}), but once a disk settles to a thin, dynamically cool state,
the formation of bars proceeds quickly. Most theoretical predictions show that
bars are long-lasting in the absence of significant galaxy interactions
\citep{ath05}, so the fraction of bars in disk galaxies can measure the
dynamical maturity of that population. For this reason, tracking the fraction
of bars in disk galaxies as a function of redshift has attracted significant
interest since the first resolved images of high redshift galaxies were
obtained \citep[\eg,][]{abr99,elm04,jog04,she08a, mel14,sim14}.

We create a sample of barred disk galaxies in GZH by applying cuts on the previous
tasks in Table~\ref{tbl:thresholds}. We first identify 11,049 ``bar
candidates,'' which are disk galaxies that are sufficiently face-on to attempt
visual identification of a bar. These galaxies were selected by applying the cuts $N_{bar} \ge 20$,
$f_{\rm features}>0.23, f_{\rm clumpy,no} > 0.30,$ and $f_{edgeon,no}>0.25$.
These galaxies are featured, non-clumpy, non-edge~on galaxies. Of these, a pure
sample of 730~barred disks was identified by applying a cut of \fbar$>0.7$. A
subsample of 50~galaxies were visually inspected and 94\% were found to contain
strong bars. A complete sample of strong and weak bars was created by applying
a cut of \fbar$>0.3$.  This sample contained 3,218~galaxies, 86\% of which were
found to contain weak or strong bars through visual inspection.

The resulting bar sample can be used to estimate the redshift evolution of bar
fraction; we find a steady decrease of \fbar$\sim 0.32$ at $z=0.4$ to
\fbar$\sim 0.24$ at $z=1.0$. This decrease agrees with
\citet{mel14}, although they report a lower overall bar fraction going from
\fbar$=0.22$ at $z=0.4$ to \fbar$=0.11$ at $z=1.0$. The difference in total bar
fraction is expected, as this analysis used a looser cut on \fbar, there is no
luminosity cut, and the use of debiased values for \ffeatures{} increases the
total amount of disks in the sample compared to \citet{mel14}. However, the
results from GZH indicate that if features are reliably identified, the bar is
always visible; for galaxies at increasing redshifts, the disks fade first but
the galactic bar remains visible \citep{mel16}. As a result, we do \emph{not}
recommend using an evolving cut on \fbar{} as a function of redshift for the
selection of barred disks.
